---
title: "IST687 Final Project - Group G"
author: "John Jefferson, Tanya MacKinnon and John O'Rourke"
date: "3/29/2022"
output:
html_document:
df_print: paged
---

This is Group G's Final Project R code.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Section 1: Initial Start - Install Packages and Load Libraries

```{r Libraries}
# Ensure packages are available
EnsurePackage<-function(x) {
  x<- as.character(x)
  
  if(!require(x,character.only = TRUE)) {
    install.packages(pkgs=x,repos="http://cran.r-project.org/")
    require(x,character.only = TRUE)
  }
}
EnsurePackage("tidyverse")
EnsurePackage("imputeTS")
EnsurePackage("ggplot2")
EnsurePackage("caret")
EnsurePackage("shiny")
EnsurePackage("corrplot")
EnsurePackage("cowplot")
EnsurePackage("lmtest")

library(lmtest)
library(tidyverse)
library(dplyr)
library(caret)
library(kernlab)
library(e1071)
library(rpart)
library(rpart.plot)
library(imputeTS)
library(ggplot2)
library(ggmap)
library(maps)
library(mapproj)
library(gridExtra)

library(corrplot)
library(cowplot)
```

## Section 2: Load data file and review data structure

```{r Importing}

#Function to import data file
importHousing <- function(filename){
  datasetHousing <- read_csv(filename)
  str(datasetHousing)
  return(datasetHousing)
}


#John O'Rouke Local file location - "F:/Syracuse/IST-687/Final Project/Dataset/housing.csv/housing.csv"
# Tanya local file location "C:/Users/tlmac/OneDrive/Documents/syracuse/ist 687/housing.csv"

#load data
file <- "F:/Syracuse/IST-687/Final Project/Dataset/housing.csv/housing.csv"
housingData <- importHousing(file)

#review data
summary(housingData)
glimpse(housingData)
sapply(housingData, sd, na.rm=TRUE)
skewness(housingData$median_income)
skewness(housingData$median_house_value)
skewness(housingData$total_bedrooms,na.rm=TRUE)
skewness(housingData$total_rooms,na.rm=TRUE)
skewness(housingData$population,na.rm=TRUE)
skewness(housingData$housing_median_age,na.rm=TRUE)
skewness(housingData$households,na.rm=TRUE)
```

## Section 3: Data Cleaning - Null Values, Factors, Unit adjustments

```{r Cleaning}

#Replace null values in total bedrooms with na_interpolation
sapply(housingData, function(x) sum(is.na(x)))
housingData$total_bedrooms <- na_interpolation(housingData$total_bedrooms)
sapply(housingData, function(x) sum(is.na(x)))

#create a factor for ocean_proximity variable
housingData %>% group_by(ocean_proximity) %>%
                summarise(total_valeu = sum(median_house_value),
                .groups = 'drop')
housingData$ocean_factor <- as.factor(housingData$ocean_proximity)
housingData %>% group_by(ocean_factor) %>%
                summarise(total_valeu = sum(median_house_value),
                .groups = 'drop')


#create new variables for adjusted units on total bedrooms, total rooms, population by houstehold
housingData$rooms_per_house= housingData$total_rooms/housingData$households
housingData$bedrooms_per_house=housingData$total_bedrooms/housingData$households
housingData$population_per_house=housingData$population/housingData$households

#change unit $10000 for median house price

housingData$median_house_value <- housingData$median_house_value/10000


#review data 
summary(housingData)
```

## Section 4: Data Exploration - Variables - histograms and boxplots

```{r exploration}

#histograms

histFun <- function(x){
  ggplot(data=housingData,aes({{x}})) + 
   geom_histogram(color = "#000000", fill = "#0099F8", bins=30) +
   geom_vline(aes(xintercept = mean({{x}})), color = "#000000", size = 1.25) +
   geom_vline(aes(xintercept = mean({{x}}) + sd({{x}})), color = "#000000", size = 1, linetype = "dashed") +
   geom_vline(aes(xintercept = mean({{x}}) - sd({{x}})), color = "#000000", size = 1, linetype = "dashed")
}

hist_house_value <- histFun(median_house_value)
hist_income <- histFun(median_income)
hist_rooms <- histFun(rooms_per_house)
hist_ocean <- histFun(as.numeric(ocean_factor))
hist_population <- histFun(population_per_house)
hist_bedrooms <- histFun(bedrooms_per_house)
hist_household <- histFun(households)
hist_age <- histFun(housing_median_age)
hist_latitude <- histFun(latitude)
hist_longitude <- histFun(longitude)

hist_grid <- plot_grid(
   hist_house_value, hist_income, hist_rooms, hist_ocean, hist_population, hist_bedrooms, hist_household, hist_age, hist_longitude, hist_latitude,
   nrow = 5,
   ncol = 2
)

hist_grid

#boxplots

boxFun <- function(y) {
  ggplot(housingData, aes({{y}})) +
    geom_boxplot(outlier.colour="red", outlier.shape=8,outlier.size=4, fill = "#0099F8")
}

boxMedianHousingValue <- boxFun(median_house_value)
boxIncome <- boxFun(median_income)
boxTotal_rooms <- boxFun(rooms_per_house)
# boxOcean <- boxFun(as.numeric(ocean_factor))
boxPopulation <- boxFun(population_per_house)
boxTotal_bedrooms <- boxFun(bedrooms_per_house)
boxHousehold <- boxFun(households)
boxAge <- boxFun(housing_median_age)
boxLatitude <- boxFun(latitude)
boxLongitude <- boxFun(longitude)

#boxplot for ocean factor against median_house_value

boxOcean <- ggplot(housingData, aes(ocean_factor, y=median_house_value)) +
    geom_boxplot(outlier.colour="red", outlier.shape=8,outlier.size=4, fill = "#0099F8")

grid.arrange(boxMedianHousingValue, boxIncome, boxTotal_rooms, boxOcean, boxPopulation, boxTotal_bedrooms, boxHousehold, boxAge, boxLatitude, boxLongitude, ncol=2, nrow=5)


```

## Section 5: Data Exploration - Correlation

```{r exploration correlation}
#Create a correlation matrix
dat <- housingData %>%
    select(median_house_value,latitude,longitude,housing_median_age,total_rooms,total_bedrooms,population,households,median_income,
           rooms_per_house,bedrooms_per_house,population_per_house)
corelation_matrix <- cor(dat)
corelation_matrix

corrplot(cor(dat),
  method = "number",
  type = "upper",
  addCoef.col = 2,   
  number.cex = 0.8
)

```

## Section 6: Data Exploration - Linear Relationships  

```{r graphical data exporation}

#Graph function
ggplotRegression <- function (fit) {
ggplot(fit$model, aes_string(x = names(fit$model)[2], y = names(fit$model)[1])) + 
  geom_point() +
  stat_smooth(method = "lm", col = "red") +
  labs(title = paste("Adj R2 = ",signif(summary(fit)$adj.r.squared, 5),
                     "Intercept =",signif(fit$coef[[1]],5 ),
                     " Slope =",signif(fit$coef[[2]], 5),
                     " P =",signif(summary(fit)$coef[2,4], 5)))
}

#Function to fit linear models

linearRegressionMHV <- function(varindep1){
  lm(formula = paste("median_house_value", "~", varindep1), data = housingData)
}

plotPop <- ggplotRegression(linearRegressionMHV("population_per_house"))
plotRooms <- ggplotRegression(linearRegressionMHV("rooms_per_house"))
plotBedrooms <- ggplotRegression(linearRegressionMHV("bedrooms_per_house"))
plotAge <- ggplotRegression(linearRegressionMHV("housing_median_age"))
plotHouseholds <- ggplotRegression(linearRegressionMHV("households"))
plotIncome <- ggplotRegression(linearRegressionMHV("median_income"))
plotLat <- ggplotRegression(linearRegressionMHV("latitude"))
plotLong <- ggplotRegression(linearRegressionMHV("longitude"))
plotOcean <- ggplotRegression(linearRegressionMHV("ocean_factor"))


grid.arrange(plotIncome,plotRooms,plotOcean, plotPop,plotBedrooms,plotHouseholds,plotAge,plotLat,
plotLong, nrow=5, ncol=2)
```

## Section 7 : Data Exploration - Maps

```{r mapping}
CAMap<-map_data("state")

ggplot(CAMap)+geom_point(aes(x=long,y=lat))+coord_map()

CalData<-CAMap%>%filter(region=="california")

ggplot(CalData)+geom_point(aes(x=long,y=lat))+coord_map()

map.simple<-ggplot(CalData)+geom_polygon(color="black",fill="gray",aes(x=long,y=lat,group=group))+coord_map()

map.simple

map.simple+geom_point(data = housingData,aes(x=housingData$longitude,y=housingData$latitude,color=housingData$population,size=(housingData$population)))+ggtitle("Housing Population For Each Block")

map.simple+geom_point(data = housingData,aes(x=housingData$longitude,y=housingData$latitude,color=housingData$total_rooms,size=housingData$total_rooms))+ggtitle("Housing Total Rooms For Each Block")

map.simple+geom_point(data = housingData,aes(x=housingData$longitude,y=housingData$latitude,color=housingData$total_bedrooms,size=housingData$total_bedrooms))+ggtitle("Housing Total Bedrooms For Each Block")

map.simple+geom_point(data = housingData,aes(x=housingData$longitude,y=housingData$latitude,color=housingData$households,size=housingData$households))+ggtitle("Housing Households For Each Block")

map.simple+geom_point(data = housingData,aes(x=housingData$longitude,y=housingData$latitude,color=housingData$median_income,size=housingData$median_income))+ggtitle("Housing Median Income For Each Block")

map.simple+geom_point(data = housingData,aes(x=housingData$longitude,y=housingData$latitude,color=housingData$housing_median_age,size=housingData$housing_median_age))+ggtitle("Housing Median Age For Each Block")

map.simple+geom_point(data = housingData,aes(x=housingData$longitude,y=housingData$latitude,color=housingData$ocean_proximity,size=housingData$ocean_proximity))+ggtitle("Housing Ocean Promixy to a Beach For Each Block")

map.simple+geom_point(data = housingData,aes(x=housingData$longitude,y=housingData$latitude,color=housingData$median_house_value,size=housingData$median_house_value))+ggtitle("Housing Median House Value For Each Block")


```

## Section 9: Linear Model

```{r Linear Modelling}
#

housingModel_allVariables <- lm(formula = median_house_value ~ latitude + longitude + 
                     population_per_house +rooms_per_house + housing_median_age + 
                     bedrooms_per_house +households + median_income + ocean_factor, data = housingData)
summary(housingModel_allVariables)

housingModel_IncomeOcean <- lm(formula = median_house_value ~median_income + ocean_factor, data = housingData)
summary(housingModel_IncomeOcean)

housingModel_IncomeOceanRooms <- lm(formula = median_house_value ~median_income + ocean_factor + rooms_per_house, data = housingData)
summary(housingModel_IncomeOceanRooms)

housingModel_Location <- lm(formula = median_house_value ~ ocean_factor + latitude + longitude , data = housingData)
summary(housingModel_Location)

housingModel_Income <- lm(formula = median_house_value ~ median_income , data = housingData)
summary(housingModel_Income)

housingModel_Density <- lm(formula = median_house_value ~ population_per_house + bedrooms_per_house + rooms_per_house + households, data = housingData)
summary(housingModel_Density)

housingModel_ocean <- lm(formula = median_house_value ~ ocean_factor, data = housingData)
summary(housingModel_ocean)
```

```{r stepped linear model}
#

housingModel_step1 <- lm(formula = median_house_value ~ median_income + rooms_per_house + latitude + housing_median_age +
                           + households + longitude + bedrooms_per_house + population_per_house, data = housingData)
summary(housingModel_step1)

housingModel_step2 <- lm(formula = median_house_value ~ median_income + rooms_per_house + latitude + housing_median_age +
                           + households + longitude + bedrooms_per_house , data = housingData)
summary(housingModel_step2)

housingModel_step3 <- lm(formula = median_house_value ~ median_income + rooms_per_house + latitude + housing_median_age +
                           + households + longitude  , data = housingData)
summary(housingModel_step3)

housingModel_step4 <- lm(formula = median_house_value ~ median_income + rooms_per_house + latitude + housing_median_age +
                           + households  , data = housingData)
summary(housingModel_step4)

housingModel_step5 <- lm(formula = median_house_value ~ median_income + rooms_per_house + latitude + housing_median_age,                         data = housingData)
summary(housingModel_step5)

housingModel_step6 <- lm(formula = median_house_value ~ median_income + rooms_per_house + latitude,data = housingData)
summary(housingModel_step6)

housingModel_step7 <- lm(formula = median_house_value ~ median_income + rooms_per_house ,data = housingData)
summary(housingModel_step7)


```

## Section 10: Partition data to Train and Test

```{r Training and Testing set creation}
#

set.seed(127)

trainFrame <- createDataPartition(y=housingData$median_house_value, p=.6667, list = FALSE)
trainData <- housingData[trainFrame,]
testData <- housingData[-trainFrame,]
dim(trainData)
dim(testData)

```

#Section 11: Train and Test the model

```{r Initial Models}
#model 1 - all variables

kfold <- trainControl(method="repeatedcv",number=10,repeats=4)

train_all <- train(median_house_value ~ latitude + longitude + 
                    population_per_house + rooms_per_house + housing_median_age + 
                    bedrooms_per_house +  households + median_income + ocean_factor, 
                    data = trainData,
                    trControl=trainControl(method="none"),
                    method="lm")
summary(train_all)
predict_all <-  predict(train_all,testData)
errors <- testData$median_house_value - predict_all
RMSE <- sqrt(mean(errors^2))
RMSE
r2 <- cor(predict_all,testData$median_house_value)^2
r2

train_all_kfold <- train( median_house_value ~ latitude + longitude + 
                    population +total_rooms + housing_median_age + 
                    total_bedrooms +  households + median_income + ocean_factor, 
                    data = trainData,
                    trControl=kfold,
                    method="lm")

summary(train_all_kfold)
predict_all_kfold <-  predict(train_all_kfold,testData)
errors <- testData$median_house_value - predict_all_kfold
RMSE <- sqrt(mean(errors^2))
RMSE
r2 <- cor(predict_all_kfold,testData$median_house_value)^2
r2


#model 2 - income
train_income <- train(median_house_value ~  median_income, 
                    data = trainData,
                    trControl=trainControl(method="none"),
                    method="lm")
summary(train_income)
predict_income <-  predict(train_income,testData)
errors <- testData$median_house_value - predict_income
RMSE <- sqrt(mean(errors^2))
RMSE
r2 <- cor(predict_income,testData$median_house_value)^2
r2


train_income_kfold <- train( median_house_value ~ median_income, 
                    data = trainData,
                    trControl=kfold,
                    method="lm")
summary(train_income_kfold)
predict_income_kfold <-  predict(train_income_kfold,testData)
errors <- testData$median_house_value - predict_income_kfold
RMSE <- sqrt(mean(errors^2))
RMSE
r2 <- cor(predict_income_kfold,testData$median_house_value)^2
r2

#model 3 - density
train_density <- train(median_house_value ~  population_per_house + bedrooms_per_house + rooms_per_house + households, 
                    data = trainData,
                    trControl=trainControl(method="none"),
                    method="lm")
summary(train_density)
predict_density <-  predict(train_density,testData)
errors <- testData$median_house_value - predict_density
RMSE <- sqrt(mean(errors^2))
RMSE
r2 <- cor(predict_density,testData$median_house_value)^2
r2


train_density_kfold <- train( median_house_value ~ population_per_house + bedrooms_per_house + rooms_per_house + households, 
                    data = trainData,
                    trControl=kfold,
                    method="lm")
summary(train_density_kfold)
predict_density_kfold <-  predict(train_density_kfold,testData)
errors <- testData$median_house_value - predict_density_kfold
RMSE <- sqrt(mean(errors^2))
RMSE
r2 <- cor(predict_density_kfold,testData$median_house_value)^2
r2



#model 3 - location
train_location <- train(median_house_value ~  ocean_factor + longitude + latitude, 
                    data = trainData,
                    trControl=trainControl(method="none"),
                    method="lm")
summary(train_location)
predict_location <-  predict(train_location,testData)
errors <- testData$median_house_value - predict_location
RMSE <- sqrt(mean(errors^2))
RMSE
r2 <- cor(predict_location,testData$median_house_value)^2
r2


train_location_kfold <- train( median_house_value ~ ocean_factor + longitude + latitude, 
                    data = trainData,
                    trControl=kfold,
                    method="lm")
summary(train_location_kfold)
predict_location_kfold <-  predict(train_location_kfold,testData)
errors <- testData$median_house_value - predict_location_kfold
RMSE <- sqrt(mean(errors^2))
RMSE
r2 <- cor(predict_location_kfold,testData$median_house_value)^2
r2


```


```{r Initial Stepped models}

kfold <- trainControl(method="repeatedcv",number=10,repeats=4)

train_step1 <- train(median_house_value ~ median_income + rooms_per_house + latitude + housing_median_age +
                           + households + longitude + bedrooms_per_house + population_per_house, 
                    data = trainData,
                    trControl=trainControl(method="none"),
                    method="lm")
summary(train_step1)
predict_step1 <-  predict(train_step1,testData)
errors <- testData$median_house_value - predict_step1
RMSE <- sqrt(mean(errors^2))
RMSE
r2 <- cor(predict_step1,testData$median_house_value)^2
r2

train_step1_kfold <- train( median_house_value ~ median_income + rooms_per_house + latitude + housing_median_age +
                           + households + longitude + bedrooms_per_house + population_per_house, 
                    data = trainData,
                    trControl = kfold,
                    method="lm")
summary(train_step1_kfold)
predict_step1_kfold <-  predict(train_step1_kfold,testData)
errors <- testData$median_house_value - predict_step1_kfold
RMSE <- sqrt(mean(errors^2))
RMSE
r2 <- cor(predict_step1_kfold,testData$median_house_value)^2
r2

```


```{r Second round of stepped Models}

kfold <- trainControl(method="repeatedcv",number=10,repeats=4)

train_step2 <- train(median_house_value ~ median_income + rooms_per_house + latitude + housing_median_age +
                           + households + longitude + bedrooms_per_house , 
                    data = trainData,
                    trControl=trainControl(method="none"),
                    method="lm")
summary(train_step2)
predict_step2 <-  predict(train_step2,testData)
errors <- testData$median_house_value - predict_step2
RMSE <- sqrt(mean(errors^2))
RMSE
r2 <- cor(predict_step2,testData$median_house_value)^2
r2

train_step2_kfold <- train( median_house_value ~ median_income + rooms_per_house + latitude + housing_median_age +
                           + households + longitude + bedrooms_per_house , 
                    data = trainData,
                    trControl=kfold,
                    method="lm")
summary(train_step2_kfold)
predict_step2_kfold <-  predict(train_step2_kfold,testData)
errors <- testData$median_house_value - predict_step2_kfold
RMSE <- sqrt(mean(errors^2))
RMSE
r2 <- cor(predict_step2_kfold,testData$median_house_value)^2
r2

```

```{r Third round of stepped models}
kfold <- trainControl(method="repeatedcv",number=10,repeats=4)

train_step2 <- train(median_house_value ~ median_income + rooms_per_house + latitude + housing_median_age +
                           + households + longitude , 
                    data = trainData,
                    trControl=trainControl(method="none"),
                    method="lm")
summary(train_step2)
predict_step2 <-  predict(train_step2,testData)
errors <- testData$median_house_value - predict_step2
RMSE <- sqrt(mean(errors^2))
RMSE
r2 <- cor(predict_step2,testData$median_house_value)^2
r2



train_step2_kfold <- train( median_house_value ~ median_income + rooms_per_house + latitude + housing_median_age +
                           + households + longitude  , 
                    data = trainData,
                    trControl=kfold,
                    method="lm")
summary(train_step2_kfold)
predict_step2_kfold <-  predict(train_step2_kfold,testData)
errors <- testData$median_house_value - predict_step2_kfold
RMSE <- sqrt(mean(errors^2))
RMSE
r2 <- cor(predict_step2_kfold,testData$median_house_value)^2
r2
```


```{r Fourth round of stepped models}
kfold <- trainControl(method="repeatedcv",number=10,repeats=4)

train_step2 <- train(median_house_value ~ median_income + rooms_per_house + latitude + housing_median_age +households, 
                    data = trainData,
                    trControl=trainControl(method="none"),
                    method="lm")
summary(train_step2)
predict_step2 <-  predict(train_step2,testData)
errors <- testData$median_house_value - predict_step2
RMSE <- sqrt(mean(errors^2))
RMSE
r2 <- cor(predict_step2,testData$median_house_value)^2
r2

kfold <- trainControl(method="repeatedcv",number=10,repeats=4)

train_step2_kfold <- train( median_house_value ~ median_income + rooms_per_house 
                            + latitude + housing_median_age + households, 
                    data = trainData,
                    trControl=kfold,
                    method="lm")

summary(train_step2_kfold)
predict_step2_kfold <-  predict(train_step2_kfold,testData)
errors <- testData$median_house_value - predict_step2_kfold
RMSE <- sqrt(mean(errors^2))
RMSE
r2 <- cor(predict_step2_kfold,testData$median_house_value)^2
r2
```

```{r Fifth round of stepped models}
kfold <- trainControl(method="repeatedcv",number=10,repeats=4)

train_step2 <- train(median_house_value ~ median_income + rooms_per_house + latitude + housing_median_age , 
                    data = trainData,
                    trControl=trainControl(method="none"),
                    method="lm")
summary(train_step2)
predict_step2 <-  predict(train_step2,testData)
errors <- testData$median_house_value - predict_step2
RMSE <- sqrt(mean(errors^2))
RMSE
r2 <- cor(predict_step2,testData$median_house_value)^2
r2



train_step2_kfold <- train( median_house_value ~ median_income + rooms_per_house + latitude + housing_median_age , 
                    data = trainData,
                    trControl=kfold,
                    method="lm")
summary(train_step2_kfold)
predict_step2_kfold <-  predict(train_step2_kfold,testData)
errors <- testData$median_house_value - predict_step2_kfold
RMSE <- sqrt(mean(errors^2))
RMSE
r2 <- cor(predict_step2_kfold,testData$median_house_value)^2
r2
```

```{r Sixth round of stepped models}
kfold <- trainControl(method="repeatedcv",number=10,repeats=4)

train_step2 <- train(median_house_value ~ median_income + rooms_per_house + latitude , 
                    data = trainData,
                    trControl=trainControl(method="none"),
                    method="lm")
summary(train_step2)
predict_step2 <-  predict(train_step2,testData)
errors <- testData$median_house_value - predict_step2
RMSE <- sqrt(mean(errors^2))
RMSE
r2 <- cor(predict_step2,testData$median_house_value)^2
r2



train_step2_kfold <- train( median_house_value ~ median_income + rooms_per_house + latitude , 
                    data = trainData,
                    trControl=kfold,
                    method="lm")
summary(train_step2_kfold)
predict_step2_kfold <-  predict(train_step2_kfold,testData)
errors <- testData$median_house_value - predict_step2_kfold
RMSE <- sqrt(mean(errors^2))
RMSE
r2 <- cor(predict_step2_kfold,testData$median_house_value)^2
r2
```
```{r Final round of stepped models}
kfold <- trainControl(method="repeatedcv",number=10,repeats=4)

train_step2 <- train(median_house_value ~ median_income + rooms_per_house , 
                    data = trainData,
                    trControl=trainControl(method="none"),
                    method="lm")
summary(train_step2)
predict_step2 <-  predict(train_step2,testData)
errors <- testData$median_house_value - predict_step2
RMSE <- sqrt(mean(errors^2))
RMSE
r2 <- cor(predict_step2,testData$median_house_value)^2
r2



train_step2_kfold <- train( median_house_value ~ median_income + rooms_per_house , 
                    data = trainData,
                    trControl=kfold,
                    method="lm")
summary(train_step2_kfold)
predict_step2_kfold <-  predict(train_step2_kfold,testData)
errors <- testData$median_house_value - predict_step2_kfold
RMSE <- sqrt(mean(errors^2))
RMSE
r2 <- cor(predict_step2_kfold,testData$median_house_value)^2
r2
```


```{r Final Results}
#best model test for regression assumptions

housingModel_allVariables <- lm(formula = median_house_value ~ latitude + longitude + 
                     population_per_house +rooms_per_house + housing_median_age + 
                     bedrooms_per_house +households + median_income + ocean_factor, data = housingData)
summary(housingModel_allVariables)

resettest(housingModel_allVariables, power=2, type="regressor",data = natopanel)

library(car)
vif(housingModel_allVariables)

outlierTest(housingModel_allVariables)

```

## Section 12: Experimental Dataset - Removing Outliers
```{r experiment - remove outlier from rooms_per_house}

summary(housingData)
str(housingData)

Q <- quantile(housingData$rooms_per_house, probs=c(.75), na.rm = FALSE)
Q

housingRevised<- subset(housingData,  
                        housingData$rooms_per_house < quantile(housingData$rooms_per_house, probs=c(.75)) & 
                        housingData$rooms_per_house > quantile(housingData$rooms_per_house, probs=c(.25)) , na.rm = FALSE)

str(housingRevised)


boxFun2 <- function(y) {
  ggplot(housingRevised, aes({{y}})) +
    geom_boxplot(outlier.colour="red", outlier.shape=8,outlier.size=4, fill = "#0099F8")
}

boxrooms <- boxFun2(rooms_per_house)
boxrooms



set.seed(127)

trainFrame <- createDataPartition(y=housingRevised$median_house_value, p=.6667, list = FALSE)
trainRevised <- housingRevised[trainFrame,]
testRevised <- housingRevised[-trainFrame,]
dim(trainRevised)
dim(testData)


housingModel_allVariables_revised <- lm(formula = median_house_value ~ latitude + longitude + 
                     population_per_house +rooms_per_house + housing_median_age + 
                     bedrooms_per_house +households + median_income + ocean_factor, data = housingRevised)
summary(housingModel_allVariables_revised)




train_revised <- train(median_house_value ~ latitude + longitude + 
                     population_per_house +rooms_per_house + housing_median_age + 
                     bedrooms_per_house +households + median_income + ocean_factor, data = trainRevised,
                    trControl=trainControl(method="none"),
                    method="lm")
summary(train_revised)


predict <-  predict(train_revised,testRevised)
errors <- testRevised$median_house_value - predict
RMSE <- sqrt(mean(errors^2))
RMSE
r2 <- cor(predict,testRevised$median_house_value)^2
r2

```
